---
title: "Exploring Ecommerce Trends of Pakistan from 2016-2018"
author: "Zurnab SARFRAZ"
date: "09/04/2022"
output:
  html_document:
    toc_float: yes
    number_sections: yes
    toc: yes
    theme: flatly
  pdf_document:
    toc: yes
  word_document:
    toc: yes
number_sections: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

*Uploading the dataset for review*

```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
library(stringr)
library(lubridate)
library(lattice)
library(gridExtra)


pk <- read.csv("C:/Users/zurna/OneDrive/Desktop/NEOMA/Smester 2/Programing/Exam/Pk_ecommerce.csv", header = T, sep = ",")
```

*Cleaning the dataset*

```{r}
pk[22:26] <- NULL #removing last 5 blank columns


pk <- pk %>% 
  select(21,1:20) #bringing customer id column on 1st place

pk[c(9,11,14,15,8,17,18,19,20,21)] <- NULL #removing variables that are of no use

pk<- clean_names(pk)

```

*Preparing the dataset*

```{r}
# fixing the status category for dual meaning categories
pk <- pk %>%
  mutate(status = ifelse(status=='received','complete', status)) %>%
  mutate(status = ifelse(status=='closed','complete', status)) %>%
  mutate(status = ifelse(status=='exchange','complete', status)) %>%
  mutate(status = ifelse(status=='order_refunded','refund', status)) %>%
  mutate(status = ifelse(status=='paid','complete',status)) %>%
  mutate(status = ifelse(status=='cod','complete', status)) %>%
  mutate(status = ifelse(status=='pending_paypal','pending',status)) %>%
  mutate(status = ifelse(status=="",NA,status)) %>%
  mutate(status = ifelse(status=="\\N",NA, status)) %>%
  mutate(category_name_1 = ifelse(category_name_1 =='\\N',NA,category_name_1)) %>%
  mutate(category_name_1 = ifelse(category_name_1 =='',NA,category_name_1)) %>%
  mutate(status = ifelse(status=='payment_review','processing',status)) %>%
  mutate(payment_method = ifelse(payment_method == 'cashatdoorstep','cod',payment_method))
  


```

*removing unnecessary empty rows and also changing data types*

```{r}


pk <- pk %>% 
  mutate(customer_id = as.numeric(customer_id)) %>% 
  mutate(mv = str_squish(mv)) %>%
  mutate(mv = str_replace_all(mv,"[[:punct:]]", "")) %>%
  mutate(mv = as.numeric(mv)) %>%
  mutate(created_at = mdy(created_at))

pk <- pk %>%
  filter(!is.na(customer_id)) %>%
  filter(price <400000) %>% #removing the outliers to avoid bias
  filter(mv < 800000) %>%   #removing the outliers to avoid bias
  filter(!is.na(category_name_1)) #removed NAs is category
#had to remove half a million empty rows


```

*To remove any substantial outliers*

```{r}
boxplot(pk$price) 
boxplot(pk$mv)

```

*Fixing categories names under the status column*

```{r}

pk$payment_method <- gsub("^(easypay|Easypay)\\_+\\w+", 'easypay',pk$payment_method)
pk$payment_method <- gsub("^(Easypay)", 'easypay',pk$payment_method)
pk$payment_method <- gsub("^(jazz)\\w+", 'jazz mobile payment',pk$payment_method)

unique(pk$payment_method)

```

# Task

*Top Categories With Respect To Revenue Generation & Order Frequency (to figure out what were the top revenue generating categories and which categories were prone to receive more orders. I could then compare the result with today's data for corroboration)*

## Top Categories w.r.t Order Frequency

```{r}
top_categories_qty <- pk %>%
  group_by(category_name_1) %>%
  summarise(order = sum(qty_ordered)) %>%
  arrange(desc(order))

top_categories_qty
```

## Top Categories w.r.t Revenue (Pakistan's Currency)

```{r}
top_categories_revenue <- pk %>%
  group_by(category_name_1) %>%
  summarise(revenue = sum(mv)) %>%
  arrange(desc(revenue))

top_categories_revenue #in pkr
```

## Top Categories w.r.t Revenue (in \$)

### Dollar conversion function from pkr to dollars

```{r}
dollar_conv <- function(x) {
  category <- x[1]
  dollar_today <- 185 #conversion today from dollar to pkr
  revenue_dollar <- round(x[2]/dollar_today, digits = 0)
  return(as.data.frame(c(category,revenue_dollar)))
}

  
```

```{r}
top_categories_rev_dollar <- dollar_conv(top_categories_revenue)

top_categories_rev_dollar
```

# Task

*Top 5 Products in each respective category (What were the hot sellers in each respective category)*

## Order Wise

```{r}
#Top 5 products in each category with respect to order frequency
top_5_prod_vol <- pk %>%
  group_by(category_name_1,sku) %>%
  summarise(order = sum(qty_ordered)) %>%
  arrange(desc(order))
  

top_5_prod_vol <- top_5_prod_vol %>% 
  group_by(category_name_1) %>%
  arrange(desc(category_name_1)) %>%
  dplyr:: filter(row_number() <= 5) %>% 
  ungroup

top_5_prod_vol

```

## Revenue Wise (in PKR)

```{r}
#Top 5 products in each category with respect to revenue
top_5_prod_rev <- pk %>%
  group_by(category_name_1,sku) %>%
  summarise(revenue = sum(mv)) %>%
  arrange(desc(revenue))
  

top_5_prod_rev <- top_5_prod_rev %>% 
  group_by(category_name_1) %>%
  arrange(desc(category_name_1)) %>%
  dplyr:: filter(row_number() <= 5) %>% 
  ungroup

View(top_5_prod_rev)

```

# Task

*relation between payment methods and order status (to see if there's any relation between order status and payment method being used. As Pakistan is the cash economy, I want to see how was the market reacting to COD and bank/online/mobile transfers at that time)*

```{r}
x <- pk %>%
  group_by(status,payment_method) %>%
  filter(!is.na(status)) %>%
  summarise(freq = n()) %>%
  arrange(desc(freq)) %>%
  filter(status == 'complete' | status == 'canceled')

ggplot(x) + geom_bar(aes(x = payment_method, 
               y = freq), 
           stat = 'identity', fill = x$freq, legend = T) +
  labs(x='Payment Options', y= 'Order Status - Number',title = "Status of complete & canceled orders w.r.t payment method") + 
  facet_grid(~status) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12))
  

```

# Task

*Rank Categories on overall revenue generation with discounts and without discounts (if people are sensitive towards discounts or not or to what extent)*

```{r}
with_disc <- pk %>%
  filter(discount_amount>0)
  
  
no_disc <- pk %>%
  filter(discount_amount==0)
```

## Revenue with discounts

```{r}
with_disc_sales <- with_disc %>% #in PKR
  group_by(category_name_1) %>%
  summarise(with_discounts_rev = sum(mv)) %>%
  arrange(desc(with_discounts_rev))



with_disc_sales_dollar <- dollar_conv(with_disc_sales) #in dollars

```

## Revenue without discounts

```{r}
no_disc_sales <- no_disc %>% #in PKR
  group_by(category_name_1) %>%
  summarise(no_discounts_rev = sum(mv)) %>%
  arrange(desc(no_discounts_rev))
  


no_disc_sales_dollar <- dollar_conv(no_disc_sales) #in dollars

```

## Sales and Discounts Status Combined

```{r}
sales_discounts_status <- data.frame(with_disc_sales_dollar,no_disc_sales_dollar)
names(sales_discounts_status)[1] <- 'categories'
sales_discounts_status$category_name_1.1 <- NULL
sales_discounts_status 


```

## Plot of Discount Sensitivity

### plot 1

```{r}

ggplot(sales_discounts_status) + geom_point(aes(x = with_discounts_rev, 
               y = no_discounts_rev)) +
  labs(x='revenue with discount', y= 'revenue with no discount',title = "sales revenue with sensitivity to discounts status") + 
  facet_grid(~categories) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 2, size = 10))
```

### plot 2

```{r}
ggplot(sales_discounts_status) + geom_point(aes(x = with_discounts_rev, 
               y = no_discounts_rev)) +
  labs(x='revenue with discount', y= 'revenue with no discount',title = "sales revenue with sensitivity to discounts status") + 
  facet_grid(~categories) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 2, size = 10)) +
  facet_wrap(categories~., scales = "free")
```

### plot 3

```{r}
lattice::dotplot(no_discounts_rev~with_discounts_rev|categories,data=sales_discounts_status)

```

# Task

*Order frequency & Revenue for Apple and Samsung devices in 'mobile and tablet category' (which consumer electronics giant has a heavy foot in Pakistan in mobile & tablet category as per the given data. This would also indicate the socio economic situation of online buyers. i.e. relevant brands: Apple, Samsung, Huawei, HTC, Infinix, Xiaomi, Q Mobile)*

## Hot Seller Mobiles W.r.t volume of orders

```{r}
hot_sellers_mobile_vol <- pk %>%
  filter(category_name_1 == 'Mobiles & Tablets') %>%
  select(5,8,7,11) %>% 
  filter(grepl("Apple", sku, ignore.case=TRUE) | grepl("Samsung", sku, ignore.case=TRUE) |
         grepl("Huawei", sku, ignore.case=TRUE) | grepl("HTC", sku, ignore.case=TRUE) |
         grepl("Infinix", sku, ignore.case=TRUE) | grepl("Xiaomi", sku, ignore.case=TRUE) |
         grepl("Q Mobile", sku, ignore.case=TRUE)) %>%
  group_by(sku) %>%
  summarise(orders = sum(qty_ordered)) %>%
  arrange(desc(orders)) %>%
  slice(1:20)
hot_sellers_mobile_vol
```

## Hot Seller Mobiles W.r.t revenue (in \$)

```{r}
hot_sellers_mobile_rev_pk <- pk %>%
  filter(category_name_1 == 'Mobiles & Tablets') %>%
  select(5,8,7,11) %>% 
  filter(grepl("Apple", sku, ignore.case=TRUE) | grepl("Samsung", sku, ignore.case=TRUE) |
         grepl("Huawei", sku, ignore.case=TRUE) | grepl("HTC", sku, ignore.case=TRUE) |
         grepl("Infinix", sku, ignore.case=TRUE) | grepl("Xiaomi", sku, ignore.case=TRUE) |
         grepl("Q Mobile", sku, ignore.case=TRUE)) %>%
  group_by(sku) %>%
  summarise(revenue = sum(mv)) %>%
  arrange(desc(revenue)) %>%
  slice(1:20)
 




hot_sellers_mobile_dollar <- dollar_conv(hot_sellers_mobile_rev_pk)
hot_sellers_mobile_dollar



```

## Plot of Hot Sellers w.r.t Qty & Revenue

```{r}
top_5_vol <- slice_head(hot_sellers_mobile_vol, n = 5)

p1_vol <- ggplot(top_5_vol, aes(x = sku, y = orders)) +
  geom_point() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 10)) +
  labs(x='Name of Phones', y= 'Order Frequency of Top 5',title = "Top 5 Phones w.r.t Qty Ordered")



top_5_rev <- slice_head(hot_sellers_mobile_dollar, n = 5)

p2_rev <- ggplot(top_5_rev, aes(x = sku, y = revenue)) +
  geom_point() +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 10)) +
  labs(x='Name of Phones', y= 'Revenue of Top 5',title = "Top 5 Phones w.r.t Qty Ordered")


grid.arrange(p1_vol,p2_rev, ncol=2)
```
